---
title: "Olympic Games"
format:
  dashboard: 
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/Drewskiey11/Website.git
    theme: 
    - sketchy
    - style/olympicdash.scss
logo: Images/olympics-logo.svg
logo-alt: Olympics logo with mulitcolored circles.
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(gt)
library(scales)
```

```{r}
#| label: load-data
#| message: false
olympics_full <- read_csv("data/olympics.csv")
```

```{r}
#| label: prep-data
olympics <- olympics_full |>
  filter(!is.na(medal)) |>
  separate_wider_delim(
    cols = team, 
    names = c("team", "suffix"), 
    delim = "-", 
    too_many = "merge", 
    too_few = "align_start"
  ) |>
  select(-suffix) |>
  mutate(medal = fct_relevel(medal, "Bronze", "Silver", "Gold"))
```

# 😷 Products/Protocols

```{r}
#| label: summer-prep-data
summer_olympics <- olympics |>
  filter (season == "Summer")
```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {Height="60%"}

```{r}
#| label: summer-medals-by-sport
#| title: Medals by sport
#| fig-width: 8
#| fig-height: 3
summer_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar(color = "white") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_fill_manual(
    values = c(
      "Gold" = "#d4af37",
      "Silver" = "#c0c0c0",
      "Bronze" = "#cd7f32"
    )
  ) +
  scale_x_continuous(
  breaks = seq(1000, 7000, 1000),
  labels = label_number(big.mark = ",")
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )


```

### Row - Medals by year {Height="40%"}

::: {.card title="Medals by year"}
Due to World War II, no olypmic games were held in 1940 and 1944

```{r}
#| label: summer-medals-by-year
#| fig-asp: 0.3
#| fig-width: 8
olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width="35%"}

### Row - Value boxes {height="30%"}

```{r}
#| label: summer-caculate-most-medals
summer_most_golds <- summer_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)

summer_most_silvers <- summer_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)

summer_most_bronzes <- summer_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Golds:

`{r} summer_most_golds$n`

`{r} summer_most_golds$team`
:::

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Silvers:

`{r} summer_most_silvers$n`

`{r} summer_most_silvers$team`
:::

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Bronzes:

`{r} summer_most_bronzes$n`

`{r} summer_most_bronzes$team`
:::

### Row - Tabsets of tables {.tabset height="70%"}

::: {.card title="Top 30 medals"}
Teams sorted in descending order of medals

```{r}
library(dplyr)
library(nord)
library(gt)
library(ggplot2)
#| label: summer-top30-total-medals
olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal,
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate( Team = team, Gold, Silver, Bronze, Total) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30) |>
  gt() |>
  cols_align(align = "left", columns = "Team") |>
   data_color(
     method = "numeric",
    palette = "nord::aurora"
   )
```
:::

::: {.card title="Bottom 30 medals"}
Teams sorted in ascending order of medals

```{r}
library(dplyr)
library(nord)
library(gt)
library(ggplot2)
#| label: summer-bottom30-total-medals
olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal,
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate( Team = team, Gold, Silver, Bronze, Total) |>
  arrange(Total, Team) |>
  slice_head(n = 30) |>
  gt() |>
  cols_align(align = "left", columns = "Team") |>
   data_color(
     method = "numeric",
    palette = "nord::aurora"
   )
```
:::

# 🏂 Winter Olympics

```{r}
#| label: winter-prep-data
winter_olympics <-olympics |>
  filter (season == "Winter")
```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {Height="60%"}

```{r}
#| label: winter-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
winter_olympics |>
   mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
   scale_fill_manual(
    values = c(
      "Gold" = "#d4af37",
      "Silver" = "#c0c0c0",
      "Bronze" = "#cd7f32"
    )
  ) +
  scale_x_continuous(
  breaks = seq(1000, 7000, 1000),
  labels = label_number(big.mark = ",")
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )


```

### Row - Medals by year {Height="40%"}

::: {.card title="Medals by year"}
Due to World War II, no olypmic games were held in 1940 and 1944

```{r}
#| label: winter-medals-by-year
#| fig-asp: 0.3
#| fig-width: 10
olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width="35%"}

### Row - Value boxes {height="30%"}

```{r}
#| label: winter-caculate-most-medals
winter_most_golds <-winter_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)

winter_most_silvers <-winter_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)

winter_most_bronzes <-winter_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head( n = 1)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Golds:

`{r} winter_most_golds$n`

`{r} winter_most_golds$team`
:::

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Silvers:

`{r} winter_most_silvers$n`

`{r} winter_most_silvers$team`
:::

::: {.valuebox icon="award-fill" color="#d4af37"}
Most Bronzes:

`{r} winter_most_bronzes$n`

`{r} winter_most_bronzes$team`
:::

### Row - Tabsets of tables {.tabset height="70%"}

::: {.card title="Top 30 medals"}
Teams sorted in descending order of medals

```{r}
library(dplyr)
library(nord)
library(gt)
library(ggplot2)
#| label: winter-top30-total-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal,
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate( Team = team, Gold, Silver, Bronze, Total) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30) |>
  gt() |>
  cols_align(align = "left", columns = "Team") |>
   data_color(
     method = "numeric",
    palette = "nord::aurora"
   )
```
:::

::: {.card title="Bottom 30 medals"}
Teams sorted in ascending order of medals

```{r}
library(dplyr)
library(nord)
library(gt)
library(ggplot2)
#| label: winter-bottom30-total-medals
olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal,
    values_from = n, 
    values_fill = 0
  ) |>
  mutate(Total = Bronze + Gold + Silver) |>
  relocate( Team = team, Gold, Silver, Bronze, Total) |>
  arrange(Total, Team) |>
  slice_head(n = 30) |>
  gt() |>
  cols_align(align = "left", columns = "Team") |>
   data_color(
     method = "numeric",
    palette = "nord::aurora"
   )
```
:::

# Protocals

# Coffee shops

# Covid
